---
title: "Course project 1"
author: "Diana López"
date: "3 de febrero de 2019"
output: html_document
---

## My course project 1: Reproducible research

1. Code for reading in the dataset and/or processing the data

```{r setup,}
knitr::opts_chunk$set(echo = TRUE)

activity <- read.csv("./activity.csv")
dim(activity)
names(activity)
head(activity)
str(activity)
#total number of missing data
sum(is.na(activity$steps))/dim(activity)[[1]]
#transforming the date column into date format using lubridate
library(lubridate)
activity$date<-ymd(activity$date)
length(unique(activity$date))
```

2.Histogram of the total number of steps taken each day

```{r, echo=FALSE}
colMeans(is.na(activity))
steps_day <- aggregate(steps ~ date, rm.na = TRUE, data = activity, FUN = sum)
plot(steps_day, type = "h", lwd = 10, lend = "square")
```
3. Mean and median number of steps taken each day
```{r, echo=FALSE}
aggregate(steps ~ date, data = activity, FUN = mean)
aggregate(steps ~ date, data = activity, FUN = median)
```

4. Time series plot of the average number of steps taken

```{r, echo=FALSE}
plot(aggregate(steps ~ interval, data = activity, FUN = mean), type = "l")

```

5. The 5-minute interval that, on average, contains the maximum number of steps

```{r, echo=FALSE}
max(activity$steps, na.rm = TRUE)
```

6. Code to describe and show a strategy for imputing missing data

```{r, echo=FALSE}
sum(is.na(activity))
```
7. Histogram of the total number of steps taken each day after missing values are imputed
```{r, echo=FALSE}
activity2 <- activity
sapply(activity2, class)
activity2$steps[is.na(activity2$steps)] <- mean(na.omit(activity$steps))
activity2$date <- as.Date(activity2$date, format = "%Y-%m-%d")
steps_day2 <- aggregate(steps ~ date, rm.na = TRUE, data = activity2, FUN = sum)
par(mfrow = c(1, 2))
plot(steps_day, type = "h", lwd = 5,lend = "square", main = "With NAs")
abline(h = seq(0, 20000, 2500), lty = "dashed")
plot(steps_day2, type = "h", lwd = 5, lend = "square", main = "NAs filled")
abline(h = seq(0, 20000, 2500), lty = "dashed")
```


8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r,}
library(ggplot2)
activity$date <- as.Date(strptime(activity$date, format="%Y-%m-%d"))
activity$datetype <- sapply(activity$date, function(x) {
        if (weekdays(x) == "domingo" | weekdays(x) =="sábado") 
                {y <- "Weekend"} else 
                {y <- "Weekday"}
                y
        })
activity_by_date <- aggregate(steps~interval + datetype, activity, mean, na.rm = TRUE)
plot<- ggplot(activity_by_date, aes(x = interval , y = steps, color = datetype)) +
       geom_line() +
       labs(title = "Average daily steps by type of date", x = "Interval", y = "Average number of steps") +
       facet_wrap(~datetype, ncol = 1, nrow=2)
print(plot)
```
